---
title: "Tidying Data"
author:
  - Luc-Tanton Tran, lqtran@ucdavis.edu
  - William Holley, wjholley@ucdavis.edu
output: html_document
date: "2025-02-20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(janitor)
library(gtools)
```

## The purpose of this document is to tidy data and rename variables/values for analysis and ease of understanding.

```{r}
tf <- read.csv("data/landscapegenetics/TF_pipits_2011.csv") %>% clean_names()
ps <- read.csv("data/landscapegenetics/PS_pipits_2016.csv") %>% clean_names()

tf_1 <- tf %>% select(!autocov1000m) #redefine tf as a different df
tf_ps <- merge(tf_1,ps, all=TRUE) #merge all data frames together

genomics <- read.csv("data/landscapegenetics/TF_PS_pipits.csv") %>% clean_names()
genomics_tf_ps <- merge(tf_ps, genomics, by = "sample")
```

#no more scientific notation
```{r}
options(scipen=999)
```

# Rewriting Variables

## Redefine `malaria` as "Y" or "N" (To keep consistient with Plasmodium strains)
```{r}
genomics_tf_ps <- genomics_tf_ps %>% 
  mutate(malaria = if_else(malaria == "1", "Y", "N"))
```

## Age? This one might not be necessary
```{r}
genomics_tf_ps <- genomics_tf_ps %>% 
  mutate(age = if_else(age=="A", "adult", "juvenile")) #I spelled it right this time!
```

_For now, I'll be skipping to the ps/tf df sections because it looks like a pain in the rear to code the heterozygote part. I'll come back when I learn how to make my own function._

## Straight up removing `utm` because what the hell man
```{r}
genomics_tf_ps <- genomics_tf_ps %>% 
  select(!utm)
```
~~Changing `TLR4_het` to homozygous/heterozygous~~ (Found better way)
```{r}
#genomics_tf_ps <- genomics_tf_ps %>% 
#  mutate(tlr4_het = if_else(tlr4_het == 0, "homozygous", "heterozygous"))
```
## Mutating any variable ending with `het` to "homozygous", "heterozygous"
```{r}
genomics_tf_ps <- genomics_tf_ps %>% 
  mutate(across(ends_with("het"), #if the var name ends with "het"
                ~if_else(.x==0, "homozygous", "heterozygous")))
# Check to see if it ends with 0 (homozygous). Otherwise, heterozygous.
# Should we consider shortening it to "homo" and "hetero"? Might be easier to read.
```
## Looking into making some of the dist data into categorical variables...
```{r}
quartileswater <- quantcut(genomics_tf_ps$distwater)
table(quartileswater)

quartilesurban <- quantcut(genomics_tf_ps$dist_urb)
table(quartilesurban)

quartilesfarm <- quantcut(genomics_tf_ps$distfarm)
table(quartilesfarm)

quartilespoul <- quantcut(genomics_tf_ps$distpoul)
table(quartilespoul)
```
```{r}
genomics_tf_ps <- genomics_tf_ps %>% 
  mutate(distwater_cat = case_when(distwater <= 105 ~ "close",
                                   distwater > 105 & distwater <= 258 ~ "median close",
                                   distwater > 258 & distwater <= 725 ~ "median far",
                                   distwater > 725 ~ "far"))
```

```{r}
genomics_tf_ps <- genomics_tf_ps %>% 
  mutate(disturb_cat = case_when(dist_urb <= 45 ~ "close",
                                   dist_urb > 45 & dist_urb <= 108 ~ "median close",
                                   dist_urb > 108 & dist_urb <= 282 ~ "median far",
                                   dist_urb > 282 ~ "far"))
```

```{r}
genomics_tf_ps <- genomics_tf_ps %>% 
  mutate(distfarm_cat = case_when(distfarm <= 429 ~ "close",
                                   distfarm > 429 & distfarm <= 723 ~ "median close",
                                   distfarm > 723 & distfarm <= 1240 ~ "median far",
                                   distfarm > 1240 ~ "far"))
```

```{r}
genomics_tf_ps <- genomics_tf_ps %>% 
  mutate(distpoul_cat = case_when(distpoul <= 860 ~ "close",
                                   distpoul > 860 & distpoul <= 1830 ~ "median close",
                                   distpoul> 1830 & distpoul <= 3410 ~ "median far",
                                   distpoul > 3410 ~ "far"))
```

```{r}
genomics_tf_ps %>% 
  select(sample, distwater_cat, distfarm_cat, disturb_cat, distpoul_cat)
```

## we can change the category names at some point but those were the best I could come up with in the moment
### They look good enough to me right now


## converting 
```{r}

```












## Update csv file
```{r}
write.csv(genomics_tf_ps, "data/landscapegenetics/genomics_tf_ps.csv", row.names = FALSE)
```


## Trying a different way to merge (join) datasets
```{r}
tf_ps %>% 
  left_join(genomics)
#This seems to do the same thing as our merge from above, but carries over another sample for some reason
```

